Lesson 4
========================================================

***
```{r}
setwd("~/MOOCs/Udacity/R Data Science")
pf <- read.csv("pseudo_facebook.tsv", sep='\t')
require(ggplot2)
```

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:The highest concentration of numbers of friends is for people under the age of thirty. There are also a surprisingly high number of Facebook users over the age of 90 that have high friend counts -- probably more than there actually are in reality. There also seems to be a large number of friends for a certain age over 60 but under 90.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
summary(pf$age)

ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_point() +
        xlim(13,90)
```

***

### Overplotting
Notes: Overplotting makes it difficult to tell how many points are in each region. Adding a 'alpha = 1/20' to our geom_point layer means that it will take 20 points to be the equivalent of one of the black dots in the previous plot. 

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_point(alpha=1/20) +
        xlim(13,90)
```

Notes: We can also add a jitter to our plot, because the plots seem to be lining up on top of each other, which is not a true reflection of age. If you look at the zoomed plot, the perfect vertical columns seem intuitively wrong. Jitter adds some noise to get a clearer understanding of age versus friend count.

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_jitter(alpha=1/20) +
        xlim(13,90)
```


#### What do you notice in the plot?
Response: I notice that there is an odd spike out near age 70. I also see that the majority of people with high friend counts are under age 30. But now the friend counts are not nearly as high as they were in the previous plot.

***

### Coord_trans()
Notes:

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_jitter(alpha=1/20) +
        xlim(13,90) +
        coord_trans(x = "sqrt")
```

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_point(alpha=1/20) +
        xlim(13,90) +
        coord_trans(y = "sqrt")
```


#### What do you notice?

Notes: Coord_trans changes the shape of the plot.

***

### Alpha and Jitter
Notes: Here we can add a jitter quality to a geom point instead of using geom jitter, because geom jitter cannot be layered with a coordinate transformation for sqrt against the y variable. (We can transform x, but not y.) We have to use position = position_jitter(h=0) because if we took a friendship initiated count of zero, added a jitter of noise to our point that ended up being negative, and took the sqrt of that, it could be an imaginary number, which would produce an error.

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
        geom_point(alpha=1/10, position = position_jitter(h=0)) +
        xlim(13,90) +
        coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
require(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)
```


```{r}
pf.fc_by_age <- pf %>%
        group_by(age)%>%
        summarize(friend_count_mean = mean(friend_count),
                  friend_count_median = median(friend_count),
                  n = n()) %>%
        arrange(age)

head(pf.fc_by_age, 20)
```


Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
        geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
        geom_point(alpha=1/10,
                   position = position_jitter(h=0),
                   color = "orange") +
        geom_line(stat = "summary", fun.y = mean) +
        geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.1), linetype = 2, color = "blue") +
        geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = "blue") +
        geom_line(stat = "summary", fun.y = median, color = "blue") +
        coord_cartesian(xlim = c(13,70), ylim = c(0,1000))
```

#### What are some of your observations of the plot?
Response:We can see that the mean is consistently higher than the median value -- the data gets skewed by higher outliers. We can also get a better sense of where the probable ranges are for each age group, and how "tight" the distribution happens to be.

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = "pearson")$estimate
#with(pf, cor.test(age, friend_count, method = "pearson"))$estimate
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))$estimate
```

***

### Correlation Methods
Notes: the summary statistic tells a story of a negative relationship between age and friend count. But it does not imply causation. To imply causation, we would want to run an experiment and use inferential statistics rather than inferential statistics.

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
        geom_point(alpha = 1/50,
                   position = position_jitter(h=0),
                   color = "green") +
        geom_line(stat = "summary", fun.y = mean) +
        geom_line(stat = "summary", fun.y = median, color = "red") +
        geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.1), color = "blue", linetype = 2) +
        geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.9), color = "blue", linetype = 2) +
        coord_cartesian(xlim = c(0,200), ylim = c(0,200))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
        geom_point() +
        xlim(0, quantile(pf$www_likes_received, 0.95)) +
        ylim(0, quantile(pf$likes_received, 0.95)) +
        geom_smooth(method = "lm", color = "red")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)$estimate
```

Response: Strong correlations like that can pop up when one set is actually a superset of the other. In the last example, that is what happened. The likes received on a desktop device were correlated with the total likes received, and are highly related by nature. The variables are not independent probably, so we can't really see what is driving the phenomenon, and that can help us decide which ones to not throw in together for an analysis.

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
#library(alr3)
```

Create your plot!

```{r Temp vs Month}
#data("Mitchell")
#?Mitchell
#write.csv(Mitchell, "Mitchell.csv")

Mitchell <- read.csv("Mitchell.csv")

ggplot(aes(x = Month, y = Temp), data = Mitchell) +
        geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

I am going to guess that the correlation will be approximately zero, because we are assuming that the check for correlation is looking at a linear model. Temperature data by month will be cyclical, and high temperatures should cancel out low temperatures.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)$estimate
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
        geom_point() +
        scale_x_continuous(breaks = seq(0,12*17,12))
```

***

### A New Perspective

What do you notice?
Response: There is a cyclical pattern in the data (like a sin or cosine graph).

Watch the solution video and check out the Instructor Notes!
Notes: I was right.

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- (pf$age + ((12 - pf$dob_month)/12))
```

***

### Age with Months Means

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
        group_by(age_with_months) %>%
        summarize(friend_count_mean = mean(friend_count),
                  friend_count_median = median(friend_count),
                  n = n()) %>%
        arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) +
        geom_line()
```

***

### Smoothing Conditional Means
Notes: We got two plots. One with age in years and one in age in months. The resolution is different. We have less data to estimate each conditional mean for month bins.

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age < 71)) +
        geom_line() +
        geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) +
        geom_line() +
        geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), data = subset(pf, age < 71)) +
        geom_line(stat = "summary", fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)

```

***

### Which Plot to Choose?
Notes: Sometimes you don't have to choose! We can explore the different versions. New versions don't mean that they are better. When we share work with a larger audience, one or two visualizations can be more powerful than a large portfolio of plots.

***

### Analyzing Two Variables
Reflection: I learned how to deal with generating scatter plots in R. I learned how to jitter graphs and use alpha to get a better look at the density of data points. I learned that correlation is a useful tool, but does not imply causation, nor does it capture all the finer details of what might be happening in a plot. I learned how to zoom into a plot without clipping off data. I learned how to make line graphs, change colors, and plot confidence intervals. I learned how to make data sets of finer resolution, and generate summary data for different categories we might investigate. 

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

